%\newcommand\COURSENAME{Theory of Automata, Languages, and Computation}
%\newcommand\COURSESHORTNAME{Automata}
%\newcommand\COURSENUMBER{CISS362/MATH362}

\input{myassignmentpreamble}
\input{yliow}
\input{ciss362}
\renewcommand\TITLE{Assignment 20}
\renewcommand\AUTHOR{David Campbell}
\renewcommand\EMAIL{djcampbell2@cougars.ccis.edu}

\begin{document}
\topmatter


The following is the version of the lemma that you must use:


{\bf Pumping Lemma (for regular language).}
Let $L$ be a language (over $\Sigma$).
Suppose that for each integer $n \geq 0$, there is some 
$x \in L$ such that $|x| \geq n$ and, 
if $x = uvw$ where $u,v,w \in \Sigma^*$, then
there is some integer $i_0 \geq 0$ such that
\[
uv^{i_0}w \not\in L
\]
Then $L$ is not regular. \qed

I expect all proofs to be properly written.
This means that not only must the proof be logically correct,
it must for instance be grammatically correct.
Proper mathematical notation must be used.
Make sure you check the spelling of every word.

\newpage


Q1. Textbook Q 1.29 and (d) below.

(d) $A_4 = \{ a^{42^n} \mid n \geq 0 \}$

[The answers for (a) and (c) are already given in the book.
Make sure you study it.
You only need to complete (b) and (d).]

\SOLUTION

\input{q1a.tex} \newpage

\input{q1b.tex} \newpage

\input{q1c.tex} \newpage

\input{q1d.tex} \newpage

\newpage




Q2. Textbook Q 1.30.
[This is optional and is not graded.]

\input{q2.tex}
\newpage




Q3. Textbook Q 1.46.

[I've provided the solution for (b). Make sure you study it.]

\input{q3a.tex} \newpage
\input{q3b.tex} \newpage
\input{q3c.tex} \newpage
\input{q3d.tex} \newpage


Q4. Either prove 
\[
L = \{ab^nab^n \mid n \geq 0 \}
\]
is regular by 
\underline{constructing}
a DFA or NFA or regular \underline{expression}
 for $L$
or prove it is not regular using the pumping lemma for regular languages.

[HINT: It's \lq\lq intuitively'' clear that it's
not regular.
However intuition can be used as a guide, not as a proof.
WARNING: If you're using the pumping lemma, note that 
the $uv$ might contains two types of characters, both $a$ and $b$.]

{\bf SOLUTION.}

\input{q4.tex}
\newpage

Q5. Textbook Q 1.53.

{\bf SOLUTION.}

\input{q5.tex}
\newpage


\newpage
WARNING!!! SPOILERS ON NEXT PAGE!!!

\newpage
HINT FOR 1.53:
In binary addition we have the following:
\begin{align*}
1 &= 0 + 1 \\
11 &= 0 + 11 \\
111 &= 0 + 111 \\
1111 &= 0 + 1111
\end{align*}


\end{document}
